Frozen Civil Conflicts

### 2024-03-01 paper.qmd
Draft edits

---
title: "Frozen Civil Conflicts"
subtitle: "The Role of External State Support"
author: J.L.A. Krusell
date: today
linestretch: 1.25
bibliography: library.bib
format: pdf
execute:
    cache: true
    echo: false
    warning: false
abstract: |
    Frozen conflicts are among some of the most pernicious, difficult to
    resolve conflicts. Surprisingly, little scholarly attention has been
    paid to the factors influencing their emergence. This paper addresses
    one possible dimension, namely the effect of external state support to
    conflict actors on the probability of an ongoing civil conflict
    becoming frozen. I use an original dataset coding frozen conflicts as
    an outcome among UCDP civil conflict episodes from 1975 to 2019 in
    order to construct a series of bayesian, hierarchical probit models
    motivated by an application of bargaining theory to the domain of
    frozen conflicts. Results indicate that directed support to rebels in
    the absence of similar support to the government side increases the
    probability that a civil conflict episode terminates as a frozen
    conflict. All other treatment conditions --- specifically, directed
    support provided only to the government and directed supported given
    to both sides --- are found to be inconsequential. Taken together,
    these findings help provide a more nuanced understanding of conflict
    termination beyond the common assumption that conflicts either
    conclude through a negotiated settlement or a one-sided military
    victory. Furthermore, this paper also helps shed light on the
    conditions under which actors may prefer a sub-optimal solution,
    allowing a conflict to transition into a state of unstable peace with
    an ever present threat of violent re-escalation, rather than a
    conclusive termination.
---

@@

Consider the case of the Second Libyan Civil War. In the aftermath of the overthrow of Muammar Gaddafi in 2011, central political authority fragmented and two competing governments emerged: the internationally recognized and UN sanctioned Government of National Accord (GNA) based in Tripoli and the so-called Tobruk Government comprised of the Libyan House of Representatives elected in the 2014 parliamentary elections aligned with the Libyan National Army under Khalifa Haftar. A ceasefire was signed in 2015 leaving the GNA nominally in control of Tripoli with the Tobruk government in control of the eastern portion of Libya. Intense fighting continued among other factions,

@@

As an equilibrium state

crucially, the expected utility of maintaining the dispute outweighs the potential gain from a peace agreement.

@@

```{r}
dataset <- readRDS("./data/merged_data.rds")
n_frozen <- filter(dataset, frozen == 1) |> nrow()
n_terr_frozen <- filter(dataset, frozen == 1, incompatibility == 1) |> nrow()
```

@@

```{R cache=F}
output <- system("scripts/wordcount.sh paper.qmd", intern = T)
wc <- gsub("[^\\d]", "", output, perl = T)
```

**Word Count**: `r wc`

@@

```{r}

@@

```{R}
sbc <- readRDS("./posteriors/sbc.rds") |>
    select(-.chain, -.iteration, -.draw)

@@

```{R}

>>>>

---
title: "Frozen Civil Conflicts"
subtitle: "The Role of External State Support"
authors:
    - name: J.L.A. Krusell
      affiliations:
          - name: University of Gothenburg
date: today
linestretch: 1.25
bibliography: library.bib
filters:
    - lib/luafilter/meta.lua
format:
    pdf:
        fontsize: 11pt
        papersize: A4
        include-in-header:
            - text: |
                \usepackage[font={small}]{caption}
        template-partials:
            - templates/before-body.tex
            - templates/title.tex
execute:
    cache: true
    echo: false
    warning: false
abstract: |
    Frozen conflicts are among some of the most pernicious, difficult to
    resolve conflicts. Surprisingly, little scholarly attention has been
    paid to the factors influencing their emergence. This paper addresses
    one possible dimension, namely the effect of external state support to
    conflict actors on the probability of an ongoing civil conflict
    becoming frozen. I use an original dataset coding frozen conflicts as
    an outcome among UCDP civil conflict episodes from 1975 to 2019 in
    order to construct a series of bayesian, hierarchical probit models
    motivated by an application of bargaining theory to the domain of
    frozen conflicts. Results indicate that directed support to rebels in
    the absence of similar support to the government side increases the
    probability that a civil conflict episode terminates as a frozen
    conflict. All other treatment conditions --- specifically, directed
    support provided only to the government and directed supported given
    to both sides --- are found to be inconsequential. Taken together,
    these findings help provide a more nuanced understanding of conflict
    termination beyond the common assumption that conflicts either
    conclude through a negotiated settlement or a one-sided military
    victory. Furthermore, this paper also helps shed light on the
    conditions under which actors may prefer a sub-optimal solution,
    allowing a conflict to transition into a state of unstable peace with
    an ever present threat of violent re-escalation, rather than a
    conclusive termination.
---

@@

Consider the case of the Second Libyan Civil War. In the aftermath of the overthrow of Muammar Gaddafi in 2011, central political authority fragmented and two competing governments emerged: the internationally recognized and UN sanctioned Government of National Accord (GNA) based in Tripoli and the so-called Tobruk Government comprised of the Libyan House of Representatives elected in the 2014 parliamentary elections aligned with the Libyan National Army under Khalifa Haftar. A ceasefire was signed in 2015 leaving the GNA nominally in control of Tripoli with the Tobruk government in control of the eastern portion of Libya. However, the core contested issue, legitimized control of the central government, remained unresolved with

dyadic frozen conflict

@@

Frozen conflict as an equilibrium of a situation where both actors strictly prefer a military victory but are unable to achieve a significant military advantage and are unwilling to acquiesce to a peace agreement. The belligerents will seek to minimize costs by de-escalating the conflict intensity and the conflict will eventually freeze.

Assumes that the expected utility of maintaining the dispute is greater than the utility of a peace agreement.

-  transitory war gains outweigh peaceful division of contested good
- time horizon, low discount factor (Stall war paper assumes that actors actually discount the future?). Anticipation of a future favourable shift in the balance of power
- credible commitment issues undermine payoff from a peace agreement

Unlike conflict onsets, assume that fighting has revealed information about actors' resolve and capabilities.

@@

```{r dataset}
dataset <- readRDS("./data/merged_data.rds")
n_frozen <- filter(dataset, frozen == 1) |> nrow()
n_terr_frozen <- filter(dataset, frozen == 1, incompatibility == 1) |> nrow()
```

@@

@@

```{r}
#| label: tbl-frozen

@@

```{R}
#| label: fig-sbc
#| fig-cap: |
#|     Simulation based calibration results

sbc <- readRDS("./posteriors/sbc.rds") |>
    select(-.chain, -.iteration, -.draw)

@@

```{R}
#| label: fig-robust
#| fig-cap: Robustness checks
